\Title

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.2

Another class of approaches have designed spiking neural networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron (Delorme). Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm (Bohte) and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach. However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open. A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes. One such approach is that of Lagorce, which uses the time to the last spike to encode information into so-called temporal surfaces. We have recently extended this HOTS algorithm to provide an online response and have shown performance close to that of SOTA for categorization, with a much more efficient energy budget. A series of similar algorithms are applied to real-life applications such as optical flow (Benosman), depth estimation (Dardelet), or gesture recognition (Marro). Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.

In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains.

In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}. This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion. We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.

@@

In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect. A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.

>>>>

Another class of approaches have designed spiking neural networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron (Delorme). Another type of spiking neural network attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm (Bohte) and is now widely used in methods that attempt to transfer performance from CNNs to SNNs, for example using a surrogate gradient approach. However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open. A new generation of algorithms has emerged with the advent of the so-called silicon retina, which mimics the output of the retina by representing information using spikes. In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}. This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion. We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware. A series of similar algorithms are applied to real-life applications such as optical flow (Benosman), depth estimation (Dardelet), or gesture recognition (Marro). Recently, such an algorithm has been used in the space industry which is famous for its stringent selection process.

In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking motifs~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal motifs in biological spike trains.
In neuronal models, an efficient use or detection of these spatio-temporal motifs embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect. A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.

@@
